### 2024-09-02 globals/metadata.tex
wohooh

%TODO mandatory; please add comma-separated list of keywords
\keywords{Rational series, weighted automata, polyregular function, commutative} 
\category{} %optional, e.g. invited paper
\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website

@@

\newcommand{\acknowledge}{
    I would like to thank Gaëtan Douéneau for his help and support.
}

\newcommand{\makeabstract}{
    \begin{abstract}
This paper studies which functions computed by $\Rel$-weighted automata can be
realised by $\Nat$-weighted automata, under two extra assumptions:
commutativity (the order of letters in the input does not matter) and
polynomial growth (the ouptut of the function is bounded by a polynomial in the
size of the input). This continues the work of Karhumäki, who studied the same
question under the assumption that the computed function was a polynomial,
which is more restrictive. Answering this question starts by exhibiting a
counterexample to the theorem of Karhumäki, proposing a new and valid
characterization in the case of polynomials, that lifts to the more general
setting of polynomial growth and commutativity. Surprisingly, we manage to
leverage this effective characterization to decide whether a function computed
by a commutative $\Nat$-weighted automaton of polynomial growth is star-free, a
notion borrowed from the theory of regular languages that has been the subject
of many investigations in the context of string-to-string functions
in the last decade. 
    \end{abstract}
}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CSL 2025}
\EventAcronym{ICALP}
\EventYear{2024}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{285}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%TODO mandatory; please add comma-separated list of keywords
\keywords{Rational series, weighted automata, polyregular function, commutative} 
\category{} %optional, e.g. invited paper
\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
\relatedversiondetails[linktext={arXiv:2404.02232}, cite={lopez2024com:arxiv}]{Full Version}{https://arxiv.org/abs/2404.02232}

@@

\newcommand{\acknowledge}{
}

\newcommand{\makeabstract}{
    \begin{abstract}
This paper studies which functions computed by $\Rel$-weighted automata can be
realised by $\Nat$-weighted automata, under two extra assumptions:
commutativity (the order of letters in the input does not matter) and
polynomial growth (the ouptut of the function is bounded by a polynomial in the
size of the input). This continues the work of Karhumäki, who studied the same
question under the assumption that the computed function was a polynomial,
which is more restrictive. Answering this question starts by exhibiting a
counterexample to the theorem of Karhumäki, proposing a new and valid
characterization in the case of polynomials, that lifts to the more general
setting of polynomial growth and commutativity. Surprisingly, we manage to
leverage this effective characterization to decide whether a function computed
by a commutative $\Nat$-weighted automaton of polynomial growth is star-free, a
notion borrowed from the theory of regular languages that has been the subject
of many investigations in the context of string-to-string functions
during the last decade. 
    \end{abstract}
}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{Event Short Title}
\EventAcronym{Event Acronym}
\EventYear{2025}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{285}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-09-02 parts/appendix.tex
wohooh

\begin{proofof}[decompose-polynomial:lem]
    Let us first write $f \in \ZPoly$ as a linear combination 
    $f = f_+ - f_-$ where both $f_+$ and $f_-$ are in $\NPoly$.
    We 
    compute $\omega_+$ (resp. $\omega_-$)
    by
    applying
    \cref{n-poly-combinatorics:lem} 
    to $f_+$ (resp. to $f_-$).
    Finally, let $\omega \defined \ppcm(\omega_+, \omega_-)$.

    Using this $\omega$,
    for any choice of
    $r \defined (r_1, \dots, r_n) \in \set{0, \dots, \omega - 1}$,
    there exists two \kl{natural binomial functions} $F_r^+$ and $F_r^-$
    associated to $f_+$ and $f_-$,
    such that $X_1, \dots, X_p \geq 1$:
    \begin{equation*}
        f\left(
        a_1^{\omega X_1 + r_1} \cdots a_n ^{\omega X_n + r_n}\right) 
        = F_r^+(X_1, \dots, X_p) - F_r^-(X_1, \dots, X_p) \quad .
    \end{equation*}
    Now, for large enough values of $x_1, \dots, x_p$,
    this means that 
    $f\left(
        a_1^{\omega X_1 + r_1} \cdots a_n ^{\omega X_n + r_n}\right)$
    coincides with an \kl{integer binomial polynomial},
    and we can therefore conclude selecting different values of $r$ and
    setting some subsets of variables to be assigned to the constant $0$.
\end{proofof}

>>>>

\begin{proofof}[decompose-polynomial:lem]
    Let us first write $f \in \ZPoly$ as a linear combination 
    $f = f_+ - f_-$ where both $f_+$ and $f_-$ are in $\NPoly$.
    We 
    compute $\omega_+$ (resp. $\omega_-$)
    by
    applying
    \cref{n-poly-combinatorics:lem} 
    to $f_+$ (resp. to $f_-$).
    Finally, let $\omega$ be the least common multiple
    of $\omega_+$ and $\omega_-$, i.e. $\omega \defined \ppcm(\omega_+, \omega_-)$.

    Using this $\omega$,
    for any choice of
    $r \defined (r_1, \dots, r_n) \in \set{0, \dots, \omega - 1}$,
    there exists two \kl{natural binomial functions} $F_r^+$ and $F_r^-$
    associated to $f_+$ and $f_-$,
    such that $x_1, \dots, x_p \geq 1$:
    \begin{equation*}
        f\left(
        a_1^{\omega x_1 + r_1} \cdots a_n ^{\omega x_n + r_n}\right) 
        = F_r^+(x_1, \dots, x_p) - F_r^-(x_1, \dots, x_p) \quad .
    \end{equation*}
    Now, for large enough values of $x_1, \dots, x_p$,
    this means that 
    $f\left(
        a_1^{\omega x_1 + r_1} \cdots a_n ^{\omega x_n + r_n}\right)$
    coincides with an \kl{integer binomial polynomial},
    and we can therefore conclude selecting different values of $r$ and
    setting some subsets of variables to be assigned to the constant $0$.
\end{proofof}

### 2024-09-02 parts/correction.tex
wohooh

%! TeX program = xelatex
%! lang = en-US
\subsection{The Corrected Theorem}
\label{sec:proof}

\AP
The counterexample provided by \cref{def:bad-polynomial} relies on the fact
that $\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-polyregular functions} are.
Indeed, the polynomial $\BadPoly$ introduced in
\cref{def:bad-polynomial} satisfies $\BadPoly(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$.
In this section, we prove that closing
$\CoveredPoly$ under variable assignments is enough to recover from
\cref{karh:thm}.
We use the following notation to fix the value of some indeterminate, if
$P(X,Y)$ is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is
the polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial
function from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial
	\Nat$, the restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with
indeterminates $\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the
variables of the domain of $\nu$.

@@

\subsubsection{From $\Nat$-polyregular functions to polynomials}

\AP Let us now prove that \kl{$\Nat$-rational polynomials} are in
$\CorrectPoly$. This follows from the correct implication in the statement of
\cref{karh:thm}, but we provide a self-contained proof using a refinement of
the classical combinatorial arguments for $\ZPoly$ \cite[Lemma 4.16]{CDTL23}
and $\NPoly$ \cite[Lemma 5.37]{DOUE23}. We take extra care to reprove in our
upcoming \cref{n-poly-combinatorics:lem} a finer statement that has two main
goals: first, it formulates the traditional \emph{pumping argument} as a
composition property using \kl{(noncommutative output) polynomial functions},
allowing us to use the same formalism in the upcoming
\cref{k-combinatorial:def,ultimately-polynomial:def} and highlight the role of
\kl{commutative} \kl{polyregular functions} in the broader study of
\kl{polyregular functions}; then we also give a precise shape of the
polynomials that arise from such \emph{pumping arguments}, which was lacking in
former similar statements where $\Rat[X_1, \dots, X_n]$ was used as an over
approximation.

@@

\AP Remark that our second goal is at odds with our restriction to polynomials
with coefficients in $\Rel$, because of the polynomial $X(X-1)/2$
(\cref{q-polynomial-n-poly:ex}). However, we argue that up to a suitable change
of basis for the ring of polynomials, all coefficients will actually be in
$\Nat$. This relatively simple trick works by considering as ``basic
polynomials" the \intro{binomial function} of the form $\binom{X - \ell}{k}$,
where $\ell$ and $k$ are natural numbers. We extend this to \intro{natural
	binomial functions} that are obtained by considering $\Nat$-linear combinations
of products of \kl{binomial functions}, that is, we consider functions that
have the following shape: $f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i \prod_{j =
		1}^k \binom{x_j - p_{i,j}}{k_{i,j}}$. Beware that $\binom{X - \ell}{k}$ is
defined to be $0$ when $X \leq \ell$, and is therefore not a polynomial.

@@

\begin{lemma}
	\label{n-sf-combinatorics:lem}
	Let $f$ be a \kl{star-free $\Nat$-polyregular} function.
	There exists a computable $s \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(X_1 + s, \dots, X_p + s)
		=
		F
		\quad
		\text{ over } (\Nat)^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
	We redo the proof of \cref{n-poly-combinatorics:lem}, with the extra
	assumption that the monoid $M$ is \kl(monoid){aperiodic}, hence that there
	exists $s \in \Nat$ such that $x^{s+1} = x^s$ for all $x \in M$. As a
	consequence, we can define $\omega = 1$, and
	translate inputs by $s$, instead of multiplying them by $\omega$.
\end{proof}

@@

\subsubsection{From polynomials to $\Nat$-polyregular functions}

@@

\subsubsection{From $\Rel$ to $\Rat$}

\AP Let us conclude this section by considering polynomials with coefficients
in $\Rat$, and justify that all the combinatorial work has already happened in
$\Rel$ and $\Nat$. A first observation is that from
\cref{n-poly-combinatorics:lem,n-sf-combinatorics:lem},
we know that the functions that can be computed by \kl{star-free
$\Nat$-polyregular functions} are going to be suitable combinations of
\kl{natural binomial functions}. This motivates the following definition of the
``polynomial counterpart" of a binomial coefficient, given two numbers $\ell,k \in \Nat$,
$\intro*\pbinom{X -
\ell}{k}$ defined as $(X - \ell) \cdots (X - \ell - k) / k!$,\footnote{
    In particular, $\pbinom{X - \ell}{k}$ is defined to be $1$ when $k = 0$, 
    and $X - \ell$ when $k = 1$.
} that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial polynomials}
as $\Nat$-linear combinations of products of \kl{binomial monomials}, i.e., of
the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j
- p_{i,j}}{k_{i,j}}$. Similarly, we introduce the class of \intro{integer
binomial polynomials}, which are obtained by $\Rel$-linear combinations of
products of \kl{binomial monomials}.

@@

As an immediate corollary, we completely characterize the class of
polynomials with rational coefficients that can be computed by
\kl{star-free $\Nat$-polyregular functions}.

\begin{corollary}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial},
		\item \label{int-bin-1:item} $P$ \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-2:item} $P$ \kl{represented} by a \kl{$\Rel$-polyregular function},
		\item \label{int-bin-3:item} $P(\Rel^k) \subseteq \Rel$.
	\end{enumerate}
	These properties are furthermore decidable.
\end{corollary}
\begin{proof}
	Let us first note that implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

Obtaining an analogue of \cref{integer-binomial-polynomials:cor}
for \kl{$\Nat$-polyregular functions} is more challenging. The
naïve approach would be to consider some variation around \kl{natural binomial
	polynomials}. A first observation is that the class of products of the form
$\prod_{i = 1}^n \pbinom{X_i - p_i}{k_i}$ is a generating family for
$\Rat[\vec{X}]$ but is not linearly independent. As a consequence, there may be
several ways to write a given polynomial, and as illustrated by
\cref{natural-binomial-polynomial-basis:ex}, some of them will
exhibit negative coefficients.\footnote{This was not a problem for \kl{integer
		binomial polynomials}, for which the change of basis mapped coefficients in
	$\Rel$ to coefficients $\Rel$.} Furthermore, there are some \kl{natural
	binomial polynomials} that take negative values on positive inputs, as
illustrated by \cref{natural-binomial-polynomial-negative:ex}.
Finally, restricting our attention to \kl{non-negative} \kl{natural binomial
	polynomials} suffers from the same kind of issues as
\cref{karh:thm},
witnessed by \cref{natural-binomial-polynomial-positive-bad:ex}.

@@

\AP To address the issues raised by the previous examples, we introduce the
notion of \intro{strongly natural binomial polynomials}, as the polynomials $P
\in \Rat[X]$ such that for all partial valuation $\nu \colon \topartial \Nat$,
$\restr{P}{\nu}$ is a \kl{natural binomial polynomial}. This leads to the
following characterization.

@@

>>>>

%! TeX program = xelatex
%! lang = en-US

\AP We will discuss at the end of \cref{sec:poly-to-n-poly}
why the
example of \cref{thm:counter-example} is minimal in the number of
indeterminates, which first requires us to provide a \emph{correct} analogue of
\cref{karh:thm}. The
counterexample provided by \cref{def:bad-polynomial}
relies on the fact that
$\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-rational polynomials} are. Indeed, the polynomial $\BadPoly$
introduced in \cref{def:bad-polynomial}
satisfies $\BadPoly(X,Y,1) = 3X^2 +
3Y^2 - 2XY$, which has a negative coefficient for a \kl{maximal monomial},
namely $-2XY$. In this section, we prove that closing $\CoveredPoly$ under
variable assignments is enough to recover from
\cref{karh:thm}. We
use the following notation to fix the value of some indeterminate, if $P(X,Y)$
is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is the
polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial function
from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial \Nat$, the
restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with indeterminates
$\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the variables of the
domain of $\nu$.

@@

Let us briefly remark that \cref{corrected-version:thm} is surprising given the
fact that is not possible to decide whether a polynomial $P \in \Rel[\vec{X}]$
is \kl{non-negative} (\cref{undecidable-non-negative:lem}), by
reduction to the undecidability of Hilbert's Tenth Problem
\cite{HILB1902,MATI1970}.

\begin{lemma}
    \label{undecidable-non-negative:lem}
    Let $P \in \Rel[\vec{X}]$ be a polynomial. Then, checking whether $P$ is
    \kl{non-negative} is undecidable.
\end{lemma}
\begin{proof}
    Recall that by the undecidability of the Hilbert's Tenth Problem \cite[Problem 10 page 22]{HILB1902}, it is
    undecidable whether a given polynomial $P \in \Rel[\vec{X}]$ has a solution in
    $\Rel^n$ \cite{MATI1970,DAVIS1973}.
    The result follows by observing that
    checking whether a polynomial $P \in \Rel[\vec{X}]$ has no solutions in $\Rel^n$ if and only 
    if $P^2 - 1$ is \kl{non-negative}. 
\end{proof}

The proof of \cref{corrected-version:thm} is divided into two parts. First, we
provide in \cref{sec:n-poly-to-poly}
a fine
combinatorial understanding of what functions can be computed by functions in
$\NPoly$ and $\ZPoly$. This will allow us to prove that \kl{$\Nat$-polyregular
functions} are in $\CorrectPoly$ (\cref{n-rat-correct:lem}).
Then, in \cref{sec:poly-to-n-poly}
we will
show how to compute polynomials in $\CorrectPoly$ using $\NPoly$
(\cref{lem:correct-to-n-rat}).
With this in hand, we will next generalize these results
to polynomials in $\Rat[\vec{X}]$ in \cref{sec:rel-to-rat}.

\subsection{From $\Nat$-polyregular functions to polynomials}
\label{sec:n-poly-to-poly}

\AP Let us prove that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
This fact follows from the correct implication in the statement of
\cref{karh:thm}, but
we provide a self-contained proof using a refinement of the classical
combinatorial \emph{pumping arguments} for $\ZPoly$ \cite[Lemma 4.16]{CDTL23} and
$\NPoly$ \cite[Lemma 5.37]{DOUE23}. We take extra care to reprove in our
upcoming \cref{n-poly-combinatorics:lem} a strong statement that has
two main goals. Our first goal is to highlight the role of \kl{commutative}
\kl{polyregular functions} in the broader study of \kl{polyregular functions},
which is done by reformulating the traditional pumping argument as a
composition property involving said functions, which will be reused in the
upcoming \cref{k-combinatorial:def,ultimately-polynomial:def}
in
\cref{beyond-polynomials:sec}. Our second goal
is to give a precise shape of the functions that arise from such \emph{pumping
arguments}, which was lacking in former similar statements.

@@

\AP Our second goal is achieved by understanding that \kl{$\Nat$-polyregular
functions} essentially compute binomial coefficients, as illustrated by the
polynomial $X(X-1)/2 = \binom{X}{2}$.
of \cref{q-polynomial-n-poly:ex}. A \intro{binomial function} is a
function of the form $\binom{X - \ell}{k}$, where $\ell$ and $k$ are natural
numbers. We extend this to \intro{natural binomial functions} that are obtained
by considering $\Nat$-linear combinations of products of \kl{binomial
functions}, that is, we consider functions that have the following shape:
$f(x_1, \dots, x_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \binom{x_j -
p_{i,j}}{k_{i,j}}$. Beware that $\binom{X - \ell}{k}$ is defined to be $0$ when
$X \leq \ell$, and is therefore not a polynomial. Let us immediately
prove that \kl{binomial functions}  can be \kl{represented} in $\NSF$,
generalizing \cref{q-polynomial-n-poly:ex}.

@@

\begin{lemma}
	\label{n-sf-combinatorics:lem}
	Let $f$ be a \kl{star-free $\Nat$-polyregular} function.
	There exists a computable $s \in \Nat_{\geq 1}$
	such that for all \kl{polynomial functions}
	$q \colon \Nat^p \to \Sigma^*$,
	there exists a computable \kl{natural binomial function} $F$
	such that:
	\begin{equation*}
		f \circ q(X_1 + s, \dots, X_p + s)
		=
		F
		\quad
		\text{ over } \Nat^p
		\quad .
	\end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
	We redo the proof of \cref{n-poly-combinatorics:lem}, with the extra
	assumption that the monoid $M$ is \kl(monoid){aperiodic}, hence that there
	exists $s \in \Nat$ such that $x^{s+1} = x^s$ for all $x \in M$. As a
	consequence, we can define $\omega = 1$, and
	translate inputs by $s$, instead of multiplying them by $\omega$.
\end{proof}

Because \kl{natural binomial functions} behave as polynomials with
\kl{non-negative} \kl{maximal monomials} on large enough inputs, we can
conclude from \cref{n-poly-combinatorics:lem}
that \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.

@@

\subsection{From polynomials to $\Nat$-polyregular functions}
\label{sec:poly-to-n-poly}

@@

\subsection{From $\Rel$ to $\Rat$}
\label{sec:rel-to-rat}

\AP Let us complete our analysis of polynomials \kl{represented} by $\NPoly$ or
$\ZPoly$ by considering polynomials with coefficients in $\Rat$, and justify
that all the combinatorial work has already happened in $\Rel$ and $\Nat$. A
first observation is that from \cref{n-sf-combinatorics:lem}, we
know that the polynomials that can be computed by \kl{star-free
$\Nat$-polyregular functions} are going to be suitable combinations of
\kl{natural binomial functions}. This motivates the following definition of the
``polynomial counterpart" of a binomial coefficient, given two numbers $\ell,k
\in \Nat$, $\intro*\pbinom{X - \ell}{k}$ defined as $(X - \ell) \cdots (X -
\ell - k) / k!$,\footnote{ In particular, $\pbinom{X - \ell}{k}$ is defined to
be $1$ when $k = 0$, and $X - \ell$ when $k = 1$. } that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial polynomials}
as $\Nat$-linear combinations of products of \kl{binomial monomials}, i.e., of
the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j
- p_{i,j}}{k_{i,j}}$. Similarly, we introduce the class of \intro{integer
binomial polynomials}, which are obtained by $\Rel$-linear combinations of
products of \kl{binomial monomials}.

@@

As an immediate corollary,  we completely characterize the class of polynomials
in $\Rat[\vec{X}]$ that are \kl{represented} by $\ZPoly$ as the \kl{integer
binomial polynomials}.

\begin{theorem}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial},
		\item \label{int-bin-1:item} $P$ \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-2:item} $P$ \kl{represented} by a \kl{$\Rel$-polyregular function},
        \item \label{int-bin-3:item} $P$ is \kl{integer-valued}.
	\end{enumerate}
	These properties are furthermore decidable.
\end{theorem}
\begin{proof}
	Let us first note that implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

\AP Obtaining an analogue of \cref{integer-binomial-polynomials:cor}
for \kl{$\Nat$-polyregular functions} requires a bit more work,
as polynomials in $\Rat[\vec{X}]$ that are \kl{represented} by $\NPoly$ are not
exactly \kl{natural binomial polynomials}
(\cref{natural-binomial-polynomial-negative:ex,natural-binomial-polynomial-basis:ex,natural-binomial-polynomial-positive-bad:ex}).
To address the issues raised by these examples, we introduce the notion of
\intro{strongly natural binomial polynomials}, as the polynomials $P \in
\Rat[X]$ such that for all partial valuation $\nu \colon \topartial \Nat$,
$\restr{P}{\nu}$ is a \kl{natural binomial polynomial}, which 
characterizes the class of polynomials that are \kl{represented} by $\NPoly$
(\cref{decide-rat-poly-npoly:cor}).

@@

@@

Let us remark that \cref{decide-rat-poly-npoly:cor} shows that the
class of polynomials \kl{represented} by $\NPoly$ is the same as the class of
polynomials \kl{represented} by $\NSF$, which is a non-trivial statement that
will be reused in the study of more general \kl{commutative} functions in
$\ZPoly$.

### 2024-09-02 parts/counter.tex
wohooh

%! TeX program = xelatex
%! lang = en-US
\section{$\Nat$-rational Polynomials}
\label{polynomials:sec}
%\subsection{The counterexample}
\label{sec:c-example}

\AP Before deciding $\NPoly$ inside $\ZPoly$ (which will be done in
\cref{beyond-polynomials:sec}), let us restrict our attention to simpler
functions, namely polynomials. A polynomial $P \in \Rel[X_1, \dots, X_n]$ is an
\intro{$\Nat$-rational polynomial} if it is \kl{represented} by a
\kl{$\Nat$-rational series}, in which case it is also \kl{represented} by a
\kl{$\Nat$-polyregular function}, because it has \kl{polynomial growth}
(\cref{polyregular-polynomial-growth:lemma}). It is an easy check that
polynomials with coefficients in $\Nat$ are \kl{$\Nat$-rational polynomials}
(\cref{n-poly-n-poly:example}). However, \cref{negative-but-npoly:ex} provides
a polynomial with negative coefficients that is an \kl{$\Nat$-rational
polynomial}. The analogue notion of \intro{$\Rel$-rational polynomials} is not
of particular interest, since every polynomial $P \in \Rel[X_1, \dots, X_n]$ is
\kl{represented} by a \kl{$\Rel$-polyregular function}. Our goal is to decide
if a \kl{$\Rel$-rational polynomial} (so, any polynomial) is an
\kl{$\Nat$-rational polynomial}, and if so, to compute an
\kl{$\Nat$-polyregular function} that \kl{represents} it. This particular
problem was claimed to be solved in \cite{KARH77}, using the \cref{karh:def} to
characterize \kl{$\Nat$-rational polynomials}, as restated in \cref{karh:thm}.

@@

\begin{example}
    \label{negative-not-nrat:ex}
    \label{negative-but-npoly:ex}
    \proofref{negative-not-nrat:ex}
    The polynomials $X$, $X^2 + 3$,
    and $X^2 - 2X + 2$
    are \kl{$\Nat$-rational polynomials},
    but $- X$ is a \kl{$\Rel$-rational polynomial} that is 
    not an \kl{$\Nat$-rational polynomial}.
\end{example}

@@

We will discuss at the end of \cref{sec:proof} why the example of
\cref{thm:counter-example} is minimal in the number of indeterminates, which
first requires us to provide a \emph{correct} analogue of \cref{karh:thm}.

>>>>

%! TeX program = xelatex
%! lang = en-US
\section{$\Nat$-rational Polynomials}
\label{polynomials:sec}

\AP In this section, we will completely characterize which polynomials in
$\Rat[\vec{X}]$ that are \kl{represented} by \kl{$\Nat$-polyregular functions}
(resp. \kl{$\Rel$-polyregular functions}). To that end, we start by
characterizing these classes for polynomials in $\Rel[\vec{X}]$.
We say that a polynomial $P \in \Rel[X_1, \dots, X_n]$ is an
\intro{$\Nat$-rational polynomial} if it is \kl{represented} by a
\kl{$\Nat$-rational series}, in which case it is also \kl{represented} by a
\kl{$\Nat$-polyregular function}, because it has \kl{polynomial growth}
(\cref{polyregular-polynomial-growth:lemma}).
The analogue notion of \intro{$\Rel$-rational polynomials} is not
of particular interest, since every polynomial $P \in \Rel[X_1, \dots, X_n]$ is
\kl{represented} by a \kl{$\Rel$-polyregular function}. 
It is an easy check that
polynomials with coefficients in $\Nat$ are \kl{$\Nat$-rational polynomials}
(\cref{n-poly-n-poly:example}). However, \cref{negative-but-npoly:ex} provides
a polynomial with negative coefficients that is an \kl{$\Nat$-rational
polynomial}. 
The problem of characterizing \kl{$\Nat$-rational polynomials} was claimed to
be solved in \cite{KARH77}, using the \cref{karh:def}
to  characterize
\kl{$\Nat$-rational polynomials}, as restated in \cref{karh:thm}.

@@

\begin{example}
    \label{negative-but-npoly:ex}
    \proofref{negative-not-nrat:ex}
    The polynomials $X$, $X^2 + 3$,
    and $X^2 - 2X + 2$
    are \kl{$\Nat$-rational polynomials},
    but $- X$ is a \kl{$\Rel$-rational polynomial} that is 
    not an \kl{$\Nat$-rational polynomial}.
\end{example}

@@
